---
title: "Bayesian statistics"
author: "Vincyane Badouard"
date: "29/06/2020"
output: pdf_document
---


```{r Packages}

library(tidyverse)
library(rstan)
rstan_options(auto_write = TRUE) #option pour ne pas recompiler à chaque fois !!! gain de temps
options(mc.cores = parallel::detectCores()) #option pour ajouter des coeurs au calcul
library(bayesplot) #visualiser la chaine de Markov
library(shinystan) #for interactive stan output visualization
library(rstanarm) #for Bayesian automatic regression modelling using stan
library(brms) #Bayesian generalized multivariate non-linear multilevel models using stan

```

# Stan time 

Data
```{r}
N = dim(standardcovariantsdata) [1] #rows number

# Response values
DivTaxo_Shann = standardcovariantsdata$DivTaxo_Shann
DivPhylo_Shann = standardcovariantsdata$DivPhylo_Shann
DivFctnal_Shann = standardcovariantsdata$DivFctnal_Shann

#standardized predictors
Silv = standardcovariantsdata$prop_ba_removed
Fire = standardcovariantsdata$prop_fire
Ced = standardcovariantsdata$prop_ced
Soil = standardcovariantsdata$prop_hydrosoil
Interac = covariantsdata$prop_ba_removed * covariantsdata$prop_fire # fire & sylv multiplication no standardised
Interac_std = (as.vector(scale(Interac))) # fire & sylv standardised multiplication
           
# Predictions 
N_pred = 100

## Create 100 no standardized values of each predictor
Silv_seq <- seq(min(covariantsdata$prop_ba_removed),max(covariantsdata$prop_ba_removed), length.out = N_pred)
Fire_seq <- seq(min(covariantsdata$prop_fire),max(covariantsdata$prop_fire), length.out = N_pred)
Ced_seq <- seq(min(covariantsdata$prop_ced),max(covariantsdata$prop_ced), length.out = N_pred)
Soil_seq <- seq(min(covariantsdata$prop_hydrosoil),max(covariantsdata$prop_hydrosoil), length.out = N_pred)
Interac_seq <- seq(min(Interac),max(Interac), length.out = N_pred)

# Sampled standardized predictors
# de max et min des variables non standardizées  
Silv_pred = (Silv_seq - mean(covariantsdata$prop_ba_removed))/sd(covariantsdata$prop_ba_removed) # standardized with INITIAL var mean and scale to not change relation with theta
Fire_pred = (Fire_seq - mean(covariantsdata$prop_fire))/sd(covariantsdata$prop_fire)
Ced_pred = (Ced_seq - mean(covariantsdata$prop_ced))/sd(covariantsdata$prop_ced)
Soil_pred = (Soil_seq - mean(covariantsdata$prop_hydrosoil))/sd(covariantsdata$prop_hydrosoil)
Interac_pred = (Interac_seq - mean(Interac))/sd(Interac)

# Fire_pred = (seq(min(Fire),max(Fire), length.out = N_pred)- mean(Fire))/sd(Fire) # standardized with original var mean and scale
# Ced_pred = (seq(min(Ced),max(Ced), length.out = N_pred)- mean(Ced))/sd(Ced) 
# Soil_pred = (seq(min(Soil),max(Soil), length.out = N_pred)- mean(Soil))/sd(Soil)
# Silv_pred = (as.vector(scale(Silv_seq))) 
# Fire_pred = (as.vector(scale(Fire_seq))) 
# Ced_pred = (as.vector(scale(Ced_seq)))
# Soil_pred = (as.vector(scale(Soil_seq)))
```

Just models compilation
```{r}
# Taxo
## Data
data_Taxo <- list(
  N = N, 
  DivTaxo_Shann = DivTaxo_Shann,
  Sylv = Silv,
  Fire = Fire,
  Ced = Ced,
  Soil = Soil #we take only the hydromorphic soil
  )
data_Taxo_interac <- list(
  N = N, #rows number
  DivTaxo_Shann = DivTaxo_Shann,
  Sylv = Silv,
  Fire = Fire,
  Ced = Ced,
  Soil = Soil, #we take only the hydromorphic soil
  Interac = Interac_std, # fire & sylv standardised multiplication
  N_pred = N_pred,
  Sylv_pred = Silv_pred,
  Fire_pred = Fire_pred, 
  Ced_pred = Ced_pred, 
  Soil_pred = Soil_pred,
  Interac_pred = Interac_pred
  )
## Models
DivTaxo_ShannM2_IDH_Ced <- stan("DivTaxo_ShannM2(Cedrela_IDH).stan", data = data_Taxo, iter = 2000) 
DivTaxo_ShannM3_IDH_Sylv <- stan("DivTaxo_ShannM3(Sylv_IDH).stan", data = data_Taxo, iter = 2000) 
DivTaxo_ShannM4_IDH_Fire <- stan("DivTaxo_ShannM4(Fire_IDH).stan", data = data_Taxo, iter = 2000) 
DivTaxo_ShannM5_interac_fire_sylv <- stan("DivTaxo_ShannM5(interac_fire-sylv).stan", data = data_Taxo_interac, iter = 2000) 

# Phylo
## Data
data_Phylo <- list(
  N = N, 
  DivPhylo_Shann = DivPhylo_Shann,
  Sylv = Silv,
  Fire = Fire,
  Ced = Ced,
  Soil = Soil)

data_Phylo_interac <- list(
  N = N, 
  DivPhylo_Shann = DivPhylo_Shann,
  Sylv = Silv,
  Fire = Fire,
  Ced = Ced,
  Soil = Soil, #we take only the hydromorphic soil
  Interac = Interac_std, # fire & sylv standardised multiplication
  N_pred = N_pred,
  Sylv_pred = Silv_pred,
  Fire_pred = Fire_pred, 
  Ced_pred = Ced_pred, 
  Soil_pred = Soil_pred,
  Interac_pred = Interac_pred
  )
## Models
DivPhylo_ShannM2_IDH_Ced <- stan("DivPhylo_ShannM2(Cedrela_IDH).stan", data = data_Phylo, iter = 2000) 
DivPhylo_ShannM3_IDH_Sylv <- stan("DivPhylo_ShannM3(Sylv_IDH).stan", data = data_Phylo, iter = 2000) 
DivPhylo_ShannM4_IDH_Fire <- stan("DivPhylo_ShannM4(Fire_IDH).stan", data = data_Phylo, iter = 2000) 
DivPhylo_ShannM5_interac_fire_sylv <- stan("DivPhylo_ShannM5(interac_fire-sylv).stan", data = data_Phylo_interac, iter = 2000) 

# Fctnal
## Data
data_Fctnal <- list(
  N = N, 
  DivFctnal_Shann = DivFctnal_Shann,
  Sylv = Silv,
  Fire = Fire,
  Ced = Ced,
  Soil = Soil)

data_Fctnal_interac <- list(
  N = N, 
  DivFctnal_Shann = DivFctnal_Shann,
  Sylv = Silv,
  Fire = Fire,
  Ced = Ced,
  Soil = Soil, #we take only the hydromorphic soil
  Interac = Interac_std, # fire & sylv standardised multiplication
  N_pred = N_pred,
  Sylv_pred = Silv_pred,
  Fire_pred = Fire_pred, 
  Ced_pred = Ced_pred, 
  Soil_pred = Soil_pred,
  Interac_pred = Interac_pred
  )

data_Fctnal_CedIDH <- list(
  N = N, 
  DivFctnal_Shann = DivFctnal_Shann,
  Sylv = Silv,
  Fire = Fire,
  Ced = Ced,
  Soil = Soil,
  N_pred = N_pred,
  Ced_pred = Ced_pred)


## Models
DivFctnal_ShannM2_IDH_Ced <- stan("DivFctnal_ShannM2(Cedrela_IDH).stan", data = data_Fctnal, iter = 2000) 
DivFctnal_ShannM3_IDH_Sylv <- stan("DivFctnal_ShannM3(Sylv_IDH).stan", data = data_Fctnal, iter = 2000) 
DivFctnal_ShannM4_IDH_Fire <- stan("DivFctnal_ShannM4(Fire_IDH).stan", data = data_Fctnal, iter = 2000) 
DivFctnal_ShannM5_interac_fire_sylv <- stan("DivFctnal_ShannM5(interac_fire-sylv).stan", data = data_Fctnal_interac, iter = 2000) 

```


## Stan for taxonomic diversity of Shannon

###Simple model :
_DivTaxo_Shann ~ lognormal(mu, sigma)_ LogNormal

```{r DivTaxo_ShannM0 : data list & compilation}

data <- list(
  N = dim(Diversitytable_taxshan) [1], 
  DivTaxo_Shann = Diversitytable_taxshan$DivTaxo_Shann
  )

DivTaxo_ShannM0modif <- stan("DivTaxo_ShannM0modif.stan", data = data, iter = 2000) #compilation
DivTaxo_ShannM0modif

#Chains
mcmc_trace(as.array(DivTaxo_ShannM0modif), #as.array = comme vecteurs
           facet_args=list(labeller=label_parsed), #to put in grec letters
           np = nuts_params(DivTaxo_ShannM0modif)) # np pour afficher la divergeance

#Parameters regressions
mcmc_pairs(as.array(DivTaxo_ShannM0modif))

#Posterior distribution 
mcmc_areas(as.array(DivTaxo_ShannM0modif), prob=0.95, pars = "theta0") # pars display some parameters

#Complete visialisation
# launch_shinystan(DivTaxo_ShannM0modif) 

```
* Chains have perfectly converged to theta0 = 35, with a likelihood = 21.3
* Parameters are independant.
* Theta0 is Normal and significant.


### Intermediate Disturbance Hypothesis (IDH)
#### Cedrela 
```{r}
DivTaxo_ShannM2_IDH_Ced

# Parameters :
# theta1 = sylviculture 
# theta2 = fire 
# theta3a = Cedrela parameter croissant part of intermediate HP 
# theta3b = Cedrela parameter descending part 
# theta4 = soil

#Chains
mcmc_trace(as.array(DivTaxo_ShannM2_IDH_Ced), #as.array = comme vecteur
           np = nuts_params(DivTaxo_ShannM2_IDH_Ced)) # np pour afficher la divergeance

#Parameters regressions
mcmc_pairs(as.array(DivTaxo_ShannM2_IDH_Ced)) 

#Boxplot
mcmc_intervals(DivTaxo_ShannM2_IDH_Ced, pars = c("theta0","theta1","theta2","theta3a", "theta3b","theta4"))
plot(DivTaxo_ShannM2_IDH_Ced)

```
Cedrela is not an "intermediate disturbance" : theta3a is negativ.

#### Sylviculture 
```{r}
DivTaxo_ShannM3_IDH_Sylv

# Parameters :
# theta1a = sylviculture parameter croissant part of intermediate HP 
# theta1b = sylviculture parameter descending part 
# theta2 = fire 
# theta3 = Cedrela parameter
# theta4 = soil

#Chains
mcmc_trace(as.array(DivTaxo_ShannM3_IDH_Sylv), #as.array = comme vecteur
           np = nuts_params(DivTaxo_ShannM3_IDH_Sylv)) # np pour afficher la divergeance

#Parameters regressions
mcmc_pairs(as.array(DivTaxo_ShannM3_IDH_Sylv)) 

#Boxplot
mcmc_intervals(DivTaxo_ShannM3_IDH_Sylv, pars = c("theta0","theta1a","theta1b","theta2","theta3", "theta4")) 

```
Taxonomic diversity *decreases at low BA-removed proportion and increases at higher* BA-removed proportion
It is not "intermediate disturbance" 

Positiv soil effect 

#### Fire 
```{r}
DivTaxo_ShannM4_IDH_Fire

# Parameters :
# theta1 = sylviculture 
# theta2a = fire parameter croissant part of intermediate HP 
# theta2b = fire parameter descending part
# theta3 = Cedrela parameter
# theta4 = soil

#Chains
mcmc_trace(as.array(DivTaxo_ShannM4_IDH_Fire), #as.array = comme vecteur
           np = nuts_params(DivTaxo_ShannM4_IDH_Fire)) # np pour afficher la divergeance

#Parameters regressions
mcmc_pairs(as.array(DivTaxo_ShannM4_IDH_Fire)) 

#Boxplot
mcmc_intervals(DivTaxo_ShannM4_IDH_Fire, pars = c("theta0","theta1","theta2a","theta2b","theta3", "theta4"))

```
Fire is not an "intermediate disturbance" : theta2b is positiv.

### Interaction Sylviculture * fire

```{r}
DivTaxo_ShannM5_interac_fire_sylv

# Parameters :
# theta1 = sylviculture 
# theta2 = fire 
# theta3 = Cedrela parameter
# theta4 = soil
# theta5 = Fire:Sylv interaction

#Chains
mcmc_trace(as.array(DivTaxo_ShannM5_interac_fire_sylv)) #as.array = comme vecteur

#Parameters regressions
mcmc_pairs(as.array(DivTaxo_ShannM5_interac_fire_sylv)) 

#Boxplot
mcmc_intervals(DivTaxo_ShannM5_interac_fire_sylv, pars = paste0("theta", 0:5), prob = 0.95) #display thetas (1to5) distribution

plot(DivTaxo_ShannM5_interac_fire_sylv)
mcmc_intervals(DivTaxo_ShannM5_interac_fire_sylv, regex_pars = "theta3", prob_outer = 0.9)
```
* No effect of sylviculture
* Very low positiv effect of fire but not relevant at 97.5%
* Strong negativ effect of Cedrela
* Positiv effect of hydromorphic soil
* Strong positiv effect of sylviculture & fire interaction
Ced > fire * sylv  > soil > feu > sylv

Likelihood = 23.56

#### Predictions with Stan

```{r}
# Outputs model extraction
Pred_DivTaxo <- as.data.frame(DivTaxo_ShannM5_interac_fire_sylv, pars = c("Pred_DivTaxo_Sylv", "Pred_DivTaxo_Fire", "Pred_DivTaxo_Ced", "Pred_DivTaxo_Soil", "Pred_DivTaxo_sylvXfire")) %>% 
     reshape2::melt() %>% #100 predictions are in several col -> put them in only one
     as.tbl() %>% 
     group_by(variable) %>% 
     summarise("mean" = mean(value), "Lower_bound_2.5" = quantile(value, 0.025), "Upper_bound_97.5" = quantile(value, 0.975)) %>% 
  mutate(variable = as.character(as.factor(variable))) %>%
  mutate(N_pred = as.character(extract_numeric(variable)))  # create a colomn with N_pred number

Pred_DivTaxo$variable <- str_replace_all(Pred_DivTaxo$variable, "[[:punct:]]", " ")#remove punctuation chr
Pred_DivTaxo$variable <- str_replace_all(Pred_DivTaxo$variable, "\\d", "") #remove digits
Pred_DivTaxo$variable <- trimws(Pred_DivTaxo$variable,"r")
Pred_DivTaxo$variable <- str_replace_all(Pred_DivTaxo$variable, "\\s", "_") #replace space by _

Pred_DivTaxo <- Pred_DivTaxo %>% 
  rename(Div_pred = variable) %>% 
mutate(covar_pred = ifelse(Div_pred == "Pred_DivTaxo_Sylv", "Silv_pred", NA)) %>%  
mutate(covar_pred = ifelse(Div_pred == "Pred_DivTaxo_Fire", "Fire_pred", covar_pred)) %>%  
mutate(covar_pred = ifelse(Div_pred == "Pred_DivTaxo_Ced", "Ced_pred", covar_pred)) %>%  
mutate(covar_pred = ifelse(Div_pred == "Pred_DivTaxo_Soil", "Soil_pred", covar_pred)) %>%  
mutate(covar_pred = ifelse(Div_pred == "Pred_DivTaxo_sylvXfire", "silvXfire_pred", covar_pred))  


# Sampled covariants table:

Covariants_pred <- data.frame(
Silv_pred = Silv_pred,
Fire_pred = Fire_pred, 
Ced_pred = Ced_pred, 
Soil_pred = Soil_pred,
Silv_pred1 = Silv_seq,
Fire_pred1 = Fire_seq,
Ced_pred1 = Ced_seq,
Soil_pred1 =  Soil_seq
) %>% 
  mutate(silvXfire_pred = Interac_pred) %>%
  mutate(silvXfire_pred1 = Interac_seq) %>%
  rownames_to_column(var= "N_pred")

Standpred <- Covariants_pred %>% 
  select(N_pred, Silv_pred, Fire_pred, Ced_pred, Soil_pred, silvXfire_pred) %>% 
  reshape2:: melt(id.vars = 'N_pred',
                 variable.name = "covar_pred",
                 value.name = "Stand_val") 
  
Nostandpred <- Covariants_pred %>% 
  select(N_pred, Silv_pred1, Fire_pred1, Ced_pred1, Soil_pred1, silvXfire_pred1) %>% 
  reshape2:: melt(id.vars = 'N_pred',
                 variable.name = "covar_pred",
                 value.name = "covar_val") %>% 
  mutate(covar_pred = recode(covar_pred, "'Silv_pred1' = 'Silv_pred'; 'Fire_pred1' = 'Fire_pred'; 'Ced_pred1' = 'Ced_pred'; 'Soil_pred1' = 'Soil_pred'; 'silvXfire_pred1' = 'silvXfire_pred'")) 
  
Covariants_pred <- Standpred %>% 
  left_join(Nostandpred, by = c('N_pred', 'covar_pred'))

Pred_covariantsdata_Taxo <- Pred_DivTaxo %>% 
  left_join(Covariants_pred, by = c("N_pred", "covar_pred"))
  


# Pretiction plots

Observed_values <- covariantsdataLongformat %>%
  dcast(plot ~ variable) %>% 
  melt(id.vars = c("plot",
                   "DivTaxo_Shann", "DivPhylo_Shann", "DivFctnal_Shann"),
                 variable.name = "covar_pred",
                 value.name = "covar_val") %>% 
  mutate(covar_pred = recode(covar_pred, "'prop_ba_removed' = 'Silv_pred'; 'prop_fire' = 'Fire_pred'; 'prop_ced' = 'Ced_pred'; 'silvXfire_pred' = 'silvXfire_pred'")) 

Taxoscales <- scale_y_continuous(limits = c(0, 60))

blank_data_taxo <- data.frame(group = c("Ced_pred", "Ced_pred", "Fire_pred", "Fire_pred", "Silv_pred", "Silv_pred", "silvXfire_pred", "silvXfire_pred", "Soil_pred", "Soil_pred"), x = c(0.1, 
    0.4, 0.1, 0.4, 0.1, 0.4, 0.1, 0.4, 0.1, 0.4), y = c(10, 
    60, 10, 60, 10, 60, 10, 60, 10, 60)) # the smallest common range. It will expand according to the data

png(file = "C:/Users/Utilisateur/Desktop/Stage M2 Côte d'Ivoire/Figures/Pretiction_plots_Taxo.png", width=738, height=512)

ggplot(Pred_covariantsdata_Taxo) +
 aes(x = covar_val, y = mean) +
  geom_ribbon(aes(ymin = Lower_bound_2.5, ymax = Upper_bound_97.5), alpha = 0.2) +
 geom_point(size = 0.4, colour = "#FF6600") +
  geom_point(data = Observed_values, aes(x =covar_val, y = DivTaxo_Shann ), colour = '#3e4a89', size = 0.4) +
  labs(x = "Predictors proportions", y = "Predicted Hill's number") +
  geom_blank(data = blank_data_taxo, aes(x = x, y = y)) +
 theme_minimal() +
 facet_wrap(vars(covar_pred), scales = "free")

dev.off()


```


## Stan for phylogenetic diversity of Shannon

_DivPhylo_Shann ~ lognormal(mu, sigma)_ LogNormal

### Intermediate Disturbance Hypothesis (IDH)
#### Cedrela 
```{r}
DivPhylo_ShannM2_IDH_Ced

# Parameters :
# theta1 = sylviculture 
# theta2 = fire 
# theta3a = Cedrela parameter croissant part of intermediate HP 
# theta3b = Cedrela parameter descending part 
# theta4 = soil

#Chains
mcmc_trace(as.array(DivPhylo_ShannM2_IDH_Ced), #as.array = comme vecteur
           np = nuts_params(DivPhylo_ShannM2_IDH_Ced)) # np pour afficher la divergeance

#Parameters regressions
mcmc_pairs(as.array(DivPhylo_ShannM2_IDH_Ced)) 

#Boxplot
mcmc_intervals(DivPhylo_ShannM2_IDH_Ced, pars = c("theta0","theta1","theta2","theta3a", "theta3b","theta4"))

```
Cedrela is not an "intermediate disturbance" : theta3a is negativ.


#### Sylviculture 
```{r}
DivPhylo_ShannM3_IDH_Sylv

# Parameters :
# theta1a = sylviculture parameter croissant part of intermediate HP 
# theta1b = sylviculture parameter descending part 
# theta2 = fire 
# theta3 = Cedrela parameter
# theta4 = soil

#Chains
mcmc_trace(as.array(DivPhylo_ShannM3_IDH_Sylv), #as.array = comme vecteur
           np = nuts_params(DivPhylo_ShannM3_IDH_Sylv)) # np pour afficher la divergeance

#Parameters regressions
mcmc_pairs(as.array(DivPhylo_ShannM3_IDH_Sylv)) 

#Boxplot
mcmc_intervals(DivPhylo_ShannM3_IDH_Sylv, pars = c("theta0","theta1a","theta1b","theta2","theta3", "theta4")) 

```
Sylviculture is not an "intermediate disturbance" : theta1a is negativ and theta1b is positiv.


#### Fire 
```{r}
DivPhylo_ShannM4_IDH_Fire

# Parameters :
# theta1 = sylviculture 
# theta2a = fire parameter croissant part of intermediate HP 
# theta2b = fire parameter descending part
# theta3 = Cedrela parameter
# theta4 = soil

#Chains
mcmc_trace(as.array(DivPhylo_ShannM4_IDH_Fire), #as.array = comme vecteur
           np = nuts_params(DivPhylo_ShannM4_IDH_Fire)) # np pour afficher la divergeance

#Parameters regressions
mcmc_pairs(as.array(DivPhylo_ShannM4_IDH_Fire)) 

#Boxplot
mcmc_intervals(DivPhylo_ShannM4_IDH_Fire, pars = c("theta0","theta1","theta2a","theta2b","theta3", "theta4"))

```
Fire is not an "intermediate disturbance" : theta2b is positiv.


### Interaction Sylviculture * fire

```{r}
DivPhylo_ShannM5_interac_fire_sylv

# Parameters :
# theta1 = sylviculture 
# theta2 = fire 
# theta3 = Cedrela parameter
# theta4 = soil
# theta5 = Fire:Sylv interaction

#Chains
mcmc_trace(as.array(DivPhylo_ShannM5_interac_fire_sylv), #as.array = comme vecteur
           np = nuts_params(DivPhylo_ShannM5_interac_fire_sylv)) # np pour afficher la divergeance

#Parameters regressions
mcmc_pairs(as.array(DivPhylo_ShannM5_interac_fire_sylv)) 

#Boxplot
mcmc_intervals(DivPhylo_ShannM5_interac_fire_sylv, pars = paste0("theta", 0:5)) #display thetas (1to5) distribution

```
* No effect of sylviculture and fire
* Strong negativ effect of Cedrela
* Positiv effect of hydromorphic soil
* Strong positiv effect of sylviculture & fire interaction
Ced > fire * sylv  > soil > sylv > feu 

#### Predictions with Stan

```{r}
# Outputs model extraction
Pred_DivPhylo <- as.data.frame(DivPhylo_ShannM5_interac_fire_sylv, pars = c("Pred_DivPhylo_Sylv", "Pred_DivPhylo_Fire", "Pred_DivPhylo_Ced", "Pred_DivPhylo_Soil", "Pred_DivPhylo_sylvXfire")) %>% 
     reshape2::melt() %>% #100 predictions are in several col -> put them in only one
     as.tbl() %>% 
     group_by(variable) %>% 
     summarise("mean" = mean(value), "Lower_bound_2.5" = quantile(value, 0.025), "Upper_bound_97.5" = quantile(value, 0.975)) %>% 
  mutate(variable = as.character(as.factor(variable))) %>%
  mutate(N_pred = as.character(extract_numeric(variable)))  # create a colomn with N_pred number

Pred_DivPhylo$variable <- str_replace_all(Pred_DivPhylo$variable, "[[:punct:]]", " ")#remove punctuation chr
Pred_DivPhylo$variable <- str_replace_all(Pred_DivPhylo$variable, "\\d", "") #remove digits
Pred_DivPhylo$variable <- trimws(Pred_DivPhylo$variable,"r")
Pred_DivPhylo$variable <- str_replace_all(Pred_DivPhylo$variable, "\\s", "_") #replace space by _

Pred_DivPhylo <- Pred_DivPhylo %>% 
  rename(Div_pred = variable) %>% 
mutate(covar_pred = ifelse(Div_pred == "Pred_DivPhylo_Sylv", "Silv_pred", NA)) %>%  
mutate(covar_pred = ifelse(Div_pred == "Pred_DivPhylo_Fire", "Fire_pred", covar_pred)) %>%  
mutate(covar_pred = ifelse(Div_pred == "Pred_DivPhylo_Ced", "Ced_pred", covar_pred)) %>%  
mutate(covar_pred = ifelse(Div_pred == "Pred_DivPhylo_Soil", "Soil_pred", covar_pred)) %>%  
mutate(covar_pred = ifelse(Div_pred == "Pred_DivPhylo_sylvXfire", "silvXfire_pred", covar_pred))  



Pred_covariantsdata_Phylo <- Pred_DivPhylo %>% 
  left_join(Covariants_pred, by = c("N_pred", "covar_pred"))

# Pretiction plots

blank_data_phylo <- data.frame(group = c("Ced_pred", "Ced_pred", "Fire_pred", "Fire_pred", "Silv_pred", "Silv_pred", "silvXfire_pred", "silvXfire_pred", "Soil_pred", "Soil_pred"), x = c(0.1, 
    0.4, 0.1, 0.4, 0.1, 0.4, 0.1, 0.4, 0.1, 0.4), y = c(5, 
    20, 5, 20, 5, 20, 5, 20, 5, 20)) # the smallest common range. It will expand according to the data


png(file = "C:/Users/Utilisateur/Desktop/Stage M2 Côte d'Ivoire/Figures/Pretiction_plots_Phylo.png", width=738, height=512)

ggplot(Pred_covariantsdata_Phylo) +
 aes(x = covar_val, y = mean) +
  geom_ribbon(aes(ymin = Lower_bound_2.5, ymax = Upper_bound_97.5), alpha = 0.2) +
 geom_point(size = 0.4, colour = "#FF6600") +
  geom_point(data = Observed_values, aes(x =covar_val, y = DivPhylo_Shann ), colour = '#3e4a89', size = 0.4) +
  labs(x = "Predictors proportions", y = "Predicted Hill's number") +
  geom_blank(data = blank_data_phylo, aes(x = x, y = y)) +
 theme_minimal() +
 facet_wrap(vars(covar_pred), scales = "free")

dev.off()

```
## Stan for functional diversity of Shannon

_DivFctnal_Shann ~ lognormal(mu, sigma)_ LogNormal

### Intermediate Disturbance Hypothesis (IDH)
#### Cedrela 
```{r}
DivFctnal_ShannM2_IDH_Ced

# Parameters :
# theta1 = sylviculture 
# theta2 = fire 
# theta3a = Cedrela parameter croissant part of intermediate HP 
# theta3b = Cedrela parameter descending part 
# theta4 = soil

#Chains
mcmc_trace(as.array(DivFctnal_ShannM2_IDH_Ced), #as.array = comme vecteur
           np = nuts_params(DivFctnal_ShannM2_IDH_Ced)) # np pour afficher la divergeance

#Parameters regressions
mcmc_pairs(as.array(DivFctnal_ShannM2_IDH_Ced)) 

#Boxplot
mcmc_intervals(DivFctnal_ShannM2_IDH_Ced, pars = c("theta0","theta1","theta2","theta3a", "theta3b","theta4"))

```
Cedrela is not an "intermediate disturbance" : theta3a is negativ.


#### Sylviculture 
```{r}
DivFctnal_ShannM3_IDH_Sylv

# Parameters :
# theta1a = sylviculture parameter croissant part of intermediate HP 
# theta1b = sylviculture parameter descending part 
# theta2 = fire 
# theta3 = Cedrela parameter
# theta4 = soil

#Chains
mcmc_trace(as.array(DivFctnal_ShannM3_IDH_Sylv), #as.array = comme vecteur
           np = nuts_params(DivFctnal_ShannM3_IDH_Sylv)) # np pour afficher la divergeance

#Parameters regressions
mcmc_pairs(as.array(DivFctnal_ShannM3_IDH_Sylv)) 

#Boxplot
mcmc_intervals(DivFctnal_ShannM3_IDH_Sylv, pars = c("theta0","theta1a","theta1b","theta2","theta3", "theta4")) 

```
Sylviculture is not an "intermediate disturbance" : theta1a is negativ and theta1b is positiv.


#### Fire 
```{r}
DivFctnal_ShannM4_IDH_Fire

# Parameters :
# theta1 = sylviculture 
# theta2a = fire parameter croissant part of intermediate HP 
# theta2b = fire parameter descending part
# theta3 = Cedrela parameter
# theta4 = soil

#Chains
mcmc_trace(as.array(DivFctnal_ShannM4_IDH_Fire), #as.array = comme vecteur
           np = nuts_params(DivFctnal_ShannM4_IDH_Fire)) # np pour afficher la divergeance

#Parameters regressions
mcmc_pairs(as.array(DivFctnal_ShannM4_IDH_Fire)) 

#Boxplot
mcmc_intervals(DivFctnal_ShannM4_IDH_Fire, pars = c("theta0","theta1","theta2a","theta2b","theta3", "theta4"))

```
Fire is not an "intermediate disturbance" : theta2b is positiv.


### Interaction Sylviculture * fire

```{r}
DivFctnal_ShannM5_interac_fire_sylv

# Parameters :
# theta1 = sylviculture 
# theta2 = fire 
# theta3 = Cedrela parameter
# theta4 = soil
# theta5 = Fire:Sylv interaction

#Chains
mcmc_trace(as.array(DivFctnal_ShannM5_interac_fire_sylv), #as.array = comme vecteur
           np = nuts_params(DivFctnal_ShannM5_interac_fire_sylv)) # np pour afficher la divergeance

#Parameters regressions
mcmc_pairs(as.array(DivFctnal_ShannM5_interac_fire_sylv)) 

#Boxplot
mcmc_intervals(DivFctnal_ShannM5_interac_fire_sylv, pars = paste0("theta", 0:5)) #display thetas (1to5) distribution

```
* No effect of fire
* Negativ effect of Cedrela and sylviculture
* Positiv effect of hydromorphic soil
* Positiv effect of sylviculture & fire interaction
Ced > fire * sylv  > sylv > soil  > feu 

#### Predictions with Stan

```{r}
# Outputs model extraction
Pred_DivFctnal <- as.data.frame(DivFctnal_ShannM5_interac_fire_sylv, pars = c("Pred_DivFctnal_Sylv", "Pred_DivFctnal_Fire", "Pred_DivFctnal_Ced", "Pred_DivFctnal_Soil", "Pred_DivFctnal_sylvXfire")) %>% 
     reshape2::melt() %>% #100 predictions are in several col -> put them in only one
     as.tbl() %>% 
     group_by(variable) %>% 
     summarise("mean" = mean(value), "Lower_bound_2.5" = quantile(value, 0.025), "Upper_bound_97.5" = quantile(value, 0.975)) %>% 
  mutate(variable = as.character(as.factor(variable))) %>%
  mutate(N_pred = as.character(extract_numeric(variable)))  # create a colomn with N_pred number

Pred_DivFctnal$variable <- str_replace_all(Pred_DivFctnal$variable, "[[:punct:]]", " ")#remove punctuation chr
Pred_DivFctnal$variable <- str_replace_all(Pred_DivFctnal$variable, "\\d", "") #remove digits
Pred_DivFctnal$variable <- trimws(Pred_DivFctnal$variable,"r")
Pred_DivFctnal$variable <- str_replace_all(Pred_DivFctnal$variable, "\\s", "_") #replace space by _

Pred_DivFctnal <- Pred_DivFctnal %>% 
  rename(Div_pred = variable) %>% 
mutate(covar_pred = ifelse(Div_pred == "Pred_DivFctnal_Sylv", "Silv_pred", NA)) %>%  
mutate(covar_pred = ifelse(Div_pred == "Pred_DivFctnal_Fire", "Fire_pred", covar_pred)) %>%  
mutate(covar_pred = ifelse(Div_pred == "Pred_DivFctnal_Ced", "Ced_pred", covar_pred)) %>%  
mutate(covar_pred = ifelse(Div_pred == "Pred_DivFctnal_Soil", "Soil_pred", covar_pred)) %>%  
mutate(covar_pred = ifelse(Div_pred == "Pred_DivFctnal_sylvXfire", "silvXfire_pred", covar_pred))  




Pred_covariantsdata_Fctnal <- Pred_DivFctnal %>% 
  left_join(Covariants_pred, by = c("N_pred", "covar_pred"))

# Pretiction plots

blank_data_Fctnal <- data.frame(group = c("Ced_pred", "Ced_pred", "Fire_pred", "Fire_pred", "Silv_pred", "Silv_pred", "silvXfire_pred", "silvXfire_pred", "Soil_pred", "Soil_pred"), x = c(0.1, 
    0.4, 0.1, 0.4, 0.1, 0.4, 0.1, 0.4, 0.1, 0.4), y = c(6, 
    11, 6, 11, 6, 11, 6, 11, 6, 11)) # the smallest common range. It will expand according to the data


png(file = "C:/Users/Utilisateur/Desktop/Stage M2 Côte d'Ivoire/Figures/Pretiction_plots_Fctnal.png", width=738, height=512)

ggplot(Pred_covariantsdata_Fctnal) +
 aes(x = covar_val, y = mean) +
  geom_ribbon(aes(ymin = Lower_bound_2.5, ymax = Upper_bound_97.5), alpha = 0.2) +
 geom_point(size = 0.4, colour = "#FF6600") +
  geom_point(data = Observed_values, aes(x =covar_val, y = DivFctnal_Shann ), colour = '#3e4a89', size = 0.4) +
  labs(x = "Predictors proportions", y = "Predicted Hill's number") +
  geom_blank(data = blank_data_Fctnal, aes(x = x, y = y)) +
 theme_minimal() +
 facet_wrap(vars(covar_pred), scales = "free")

dev.off()

```