---
title: "Functional diversity"
author: "Vincyane Badouard"
date: "21/07/2020"
output: html_document
---


# Diversity indices calculation
## Calculation of functional diversity indices

https://daijiang.name/en/2014/05/11/functional-diversity-in-r/

0) Complete missing values with *MICE*
1) Take only traits
2) Compute species mean of each trait
3) Standardize all the traits (*scale()*) because their units are different
4) Obtain a dissimylarity matrix *for quanti & quali variables* : Gower method (*gowdis()*) 
5) Create a dendrogram (*hclust()*)
  - choose the agglomeration method to compute it (UPGMA) (ou tt faire et calculer leur Norme2 (pas corrélation cophénétique)
6) Test if it's utrametric & binary or force to be
7) Create a list of species composition table of all the plots (with only the species present in the traits base or not) 
8) Create a subset of this tree for each plot according to its species composition
9)Compute functional diversity indices

```{r Packages}

library(tidyverse) 
library(mice)
library(FD)
library(phylobase)
library(entropart)

```

```{r}
Traitsbase_means <- read_delim("C:/Users/Utilisateur/Desktop/Stage M2 Côte d'Ivoire/Traitsbase_fordiversity.csv", 
    ";", escape_double = FALSE, col_types = cols(Deciduousness = col_character()), 
    locale = locale(decimal_mark = ","), 
    trim_ws = TRUE)
```


### Distance matrix & dendrogram

Obtain a distance matrix
```{r Obtain a distance matrix}

#Standardize numeric traits
Traitsbase_stand <- Traitsbase_means %>% 
  mutate_if(is.numeric,
            funs(as.vector(scale(.)))) %>% 
  column_to_rownames(., var = "ScientificName") #put species name in the column "rownames"

DistMatrix <- gowdis(Traitsbase_stand)
#Gower for quanti & quali variables

```

Create a dendrogram (utrametric)
```{r Create a dendrogram (utrametric)}


FctDendrogram <-hclust(DistMatrix, method = "average")
# UPGMA : arithmetical average
plot(FctDendrogram, main="Functional dendrogram")

UltrametricMatrix <- cophenetic(FctDendrogram)

class(FctDendrogram) #it is an object of "hclust" class
FctDendrogram <- as.phylo(FctDendrogram)#transform in "phylo class" for some following functions
is.ultrametric(FctDendrogram) #It's an ultrametric tree
is.binary.tree(FctDendrogram) #It's a binary tree
plot(FctDendrogram, main="Functional utrametric tree")

FctDendrogram_phylo4 <- as(FctDendrogram, "phylo4")#transform in "phylo4 class" for "subset" functions

```

A tree for each plot
```{r a tree for each plot}

# Create a vector with only species in the functional dendrogram :
DendogramSpecies <- Traitsbase_means$ScientificName

Data_Tene <- read_delim("C:/Users/Utilisateur/Desktop/Stage M2 Côte d'Ivoire/data_final.csv",
    ";", escape_double = FALSE, locale = locale(decimal_mark = ","),
    trim_ws = TRUE)
Data_Tene <- Data_Tene %>%
  rename(genus = genus.y) %>%
  rename(species = species.y)

# list with a table for each plot : 
datafctnaltree <- Data_Tene %>% #species composition table for each plot
  select(genus, species, family, plot) %>% 
  distinct() %>% 
  filter(!is.na(genus)) %>% #remove non-identified species
  filter(!(genus == "Cedrela")) %>% #Remove Cedrela of the response variable because it will be a covariant
  filter(!(species == "papaya")) %>%
  unite(genus, species, col = "ScientificName", sep = " ", remove = F) %>% 
  unite(genus, species, col = "Scientific_Name", sep = " ", remove = F) %>% 
  filter(ScientificName %in% DendogramSpecies)

#create a list with a table for each plot :
datafctnaltreelist <- split(datafctnaltree, datafctnaltree$plot)

datafctnaltreelist <- lapply(datafctnaltreelist, function(element) column_to_rownames(element, var = "Scientific_Name"))  #put species name in the column "rownames" for 'subset' function


#a dendrogram for each plot :
fctnalTreeperplot <- lapply(datafctnaltreelist, function(element) subset(FctDendrogram_phylo4, tips.include = element$ScientificName))
save(fctnalTreeperplot, file = "fctnalTreeperplot.Rdata")
# load("fctnalTreeperplot.Rdata")


```

Functional diversity indices computation
```{r Functional diversity indices computation}

# Abondances vector for all plots :
abondperplot <- Data_Tene %>% 
  filter(!is.na(genus)) %>% #remove non-identified species
  group_by(genus, species, plot) %>% #by species & plot
  summarise(N = n()) %>%  #abundances
  arrange(plot, desc(N)) %>% 
  ungroup(genus, species) %>% 
  filter(!(genus == "Cedrela")) %>% #Remove Cedrela of the response variable because it will be a covariant
  filter(!(species == "papaya")) %>%
  unite(genus, species, col = "ScientificName", sep = " ", remove = T)

#create a list with a table for each plot :
abondplotslist <- split(abondperplot, abondperplot$plot) #separate database under a factor


# For fctnal diversity specificly (dendrogram species = abondances vector species) :
 # Abondances vector 
abondperplotofdendro <- abondperplot %>% 
  filter(ScientificName %in% DendogramSpecies)

 #list for "rao" function (dendrogram species = abondances vector species)
abondperplotofdendrolist <- split(abondperplotofdendro, abondperplotofdendro$plot)



# 'bcPhyloDiversity' function necessites a 'NAMED VECTOR' as species abundances numeric vector (non specified in R help's function) :


# Species abundances numeric NAMED vector
Namedvector_fctnofdendro <- lapply(abondperplotofdendrolist,
                      function(element){
                         Namedvector_fctnofdendro <- element$N   # operation1
                         names(Namedvector_fctnofdendro) <- element$ScientificName # operation2
                         return(Namedvector_fctnofdendro) # préciser ce que la fonction doit renvoyer
})

fctnalTreeperplot <- lapply(fctnalTreeperplot, function(element) as(element, "phylo"))#transform in "phylo class" to compute diversity indices

# 'bcPhyloDiversity' function version :
DivFctnal_Shann <- mapply(function(X,Y) {
  bcPhyloDiversity(Ns = X,
                   Tree = Y,
                   q = 1, Normalize = T,
                   Correction = "Best", CheckArguments = T)
  }, X=Namedvector_fctnofdendro, Y=fctnalTreeperplot, SIMPLIFY = F) #SIMPLIFY = F : laisse la structure initiale de la liste (par plot dans ce cas)

DivFctnal_Shann_values <- lapply(DivFctnal_Shann, getElement, "Total")#extraction de la valeur (Total) de l'indice grace à la fonction "getElement"(='$')

# 'bcRao' (Simpson) function version :
DivFctnal_Shann_Rao <- mapply(function(X,Y) {
  bcRao(Ns = X,
        Tree = Y,
        Correction = "Lande", CheckArguments = T) #"Lande", the default value (equivalent to "Best")
  }, X = Namedvector_fctnofdendro, Y = fctnalTreeperplot, SIMPLIFY = F)

```

Put values in a table
```{r Put fctnal values in a table}
#'bcPhyloDiversity' function
Diversitytable_fctnalshan <- DivFctnal_Shann_values %>% 
  as_tibble(.) %>% 
  t() %>% #transpose romws & columns
  as.tibble(.) %>%
  rownames_to_column(var= "plot") %>% 
  rename(DivFctnal_Shann=V1)

# 'bcRao' (Simpson) function
Diversitytable_fctnalshanRao <- DivFctnal_Shann_Rao %>% 
  as_tibble(.) %>% 
  t() %>% #transpose romws & columns
  as.tibble(.) %>%
  rownames_to_column(var= "plot") %>% 
  rename(DivFctnal_Rao=V1)

summary(Diversitytable_fctnalshan) #min: 3.778 ; Max: 8.642

```
