---
title: "Data exploration (Stage M2)"
author: "Vincyane Badouard"
date: "14/07/2020"
output: html_document
---


```{r Packages}

library(tidyverse)
library(reshape2)
library(Hmisc) #rcorr
library(egg)#ggarrange
library(ggpubr)
library(corrplot)
library(RColorBrewer)

```

We have computed diversity indices at 3 different scales :
 - Taxonomic
 - Phylogenetic
 - Functional
 
Each type is calculated for each plot.  


```{r Covariants data base}
data_subplot <- read_delim("C:/Users/Utilisateur/Desktop/Stage M2 Côte d'Ivoire/data_subplot.csv", 
    ";", escape_double = FALSE, locale = locale(decimal_mark = ","), 
    trim_ws = TRUE)



covariantsdata <- data_subplot %>% 
   select(-biomass, -X1, -plot_subplot, -biomass, 
          -shannon_ind, -simpson_ind, -edge_dist_norm, -mean_dist_seedsource) %>% 
  group_by(plot) %>% 
  mutate(prop_ced = sum(prop_ced)) %>% 
  mutate(prop_soil1 = sum(prop_soil1)) %>% 
  mutate(prop_soil2 = sum(prop_soil2)) %>%
  mutate(prop_soil3 = sum(prop_soil3)) %>% 
  mutate(prop_fire = sum(prop_fire)) %>% 
  mutate(prop_ba_removed = sum(prop_ba_removed)) %>% 
  ungroup() %>%
  distinct() %>% 
  arrange(plot) %>% 
  mutate(plot = as.character(as.numeric(plot))) %>%
  left_join(Diversitytable_taxshan, by = "plot") %>% # Taxonomic diversity values
  left_join(Diversitytable_phyloshan, by = "plot") %>%   # Phylogenetic diversity values
  left_join(Diversitytable_fctnalshan, by = "plot")   # Functional diversity values


standardcovariantsdata <- covariantsdata %>% 
  mutate_at(c("prop_ced", "prop_soil1", "prop_soil2", "prop_soil3", "prop_fire", "prop_ba_removed"),
            funs(as.vector(scale(.)))) #To standardize variables to put them at the same scale and to keep theta0 as mean of the response variable.
 
```


```{r Covariants-diversity relation visualisation (Regressions)}

esquisse::esquisser()

# Wide-format to long-format -> facet_wrap/grid
covariantsdataLongformat <- melt(covariantsdata)

# Double 'melt' -> long-format -> facet_wrap/grid
covariantsdataLongformat %>%
  dcast(plot ~ variable) %>% 
  melt(id.vars = c("plot",
                   "DivTaxo_Shann", "DivPhylo_Shann", "DivFctnal_Shann"),
                 variable.name = "explanatory",
                 value.name = "explanatory_value") %>% 
  melt(id.vars = c("plot",
                   "explanatory", "explanatory_value"),
                 variable.name = "response",
                 value.name = "response_value") %>% 
 ggplot() +
 aes(x = explanatory_value, 
     y = response_value) +
  geom_point(size = 2.1) +
  labs(x = "Explanatory value", y = "Response value") +
 theme_minimal()+
  geom_smooth(method = "lm") +
  ggpubr::stat_cor()+
  facet_grid(response ~ explanatory, scales = "free")

```


```{r Variables density curve}

# Variables distribution under density curve
covariantsdataLongformat %>%
 ggplot() +
 aes(x = value) +
 geom_density(adjust = 1L, fill = "#3e4a89") + # default adjustment
 theme_minimal() +
 facet_wrap(vars(variable), scales = "free")


# Variables distribution under density curve & histogram
covariantsdataLongformat %>%
  ggplot() +
 aes(x = value) +
  geom_histogram(aes(y=..density..), bins = 30L, fill = "#3e4a89") +
 geom_density(alpha = .2, fill = "antiquewhite3") + # default adjustment
 labs(x = "Value", y = "Density") +
  theme_minimal() +
 facet_wrap(vars(variable), scales = "free")

```
  * DivTaxo_Shan, DivPhylo_Shan & prop_fire can be considered as Normal variables
  * prop_ced can be considered as a logNormal variable
  * prop_ba_removed seems to be binary : "non-exploited" & "exploited", and a small distribution in the "exploited" category, which can be neglected.
  Let's try to propose a factorial version of this variable :
  
```{r Add "Sylviculture" : a factorial version}

covariantsdata_plus1 <- covariantsdata %>% 
  mutate(Sylviculture = ifelse(prop_ba_removed == 0.000, "0", "1")) %>% 
  mutate(Sylviculture = as.factor(as.character(Sylviculture)))

covariantsdata_plus1Longformat <- covariantsdata_plus1 %>% 
  select(plot, Sylviculture) %>% 
  left_join((covariantsdataLongformat))

# Variables distribution histogram with/without sylviculture
covariantsdata_plus1Longformat %>%
 ggplot() +
 aes(x = value, fill = Sylviculture) +
 geom_histogram(adjust = 1L) + # default adjustment
 theme_minimal() +
 facet_wrap(vars(variable), scales = "free")


# Variables distribution under density curve with/without sylviculture
covariantsdata_plus1Longformat %>%
 ggplot() +
 aes(x = value, fill = Sylviculture) +
 geom_density(adjust = 1L, alpha = .6,) + # default adjustment
 theme_minimal() +
 facet_wrap(vars(variable), scales = "free")

```
Sylviculture presence influence the other variables : fire, Cedrela & soil proportions
but not the response variables.

```{r, Correlation tests}

# A quantitative traits base:
baseforcor_var <- covariantsdata %>% 
  select(-plot)
#pas besoin de tester la normalité : n>30.
CorMatrix_variables <- round(cor(baseforcor_var), digits = 2) # matrice de corrélation, avec 2 décimales

CorMatrixS_variables <- rcorr(as.matrix(baseforcor_var))
CorMatrix_variables <- CorMatrixS_variables$r #correlations
Pval_corr_variables <- CorMatrixS_variables$P #p-values

# Correlation plot:
corrplot(CorMatrix_variables, method="number", type="lower", col=brewer.pal(n=8, name="PuOr"), tl.col="black", tl.srt=25)
# , p.mat = Pval_corr_variables, sig.level = 0.05) #with crosses for p-values > 0.05.

```

```{r Regressions with/without sylviculture}

# Double 'melt' -> long-format -> facet_wrap/grid
covariantsdata_plus1Longformat %>%
  dcast(plot + Sylviculture ~ variable) %>% 
  melt(id.vars = c("plot", "Sylviculture", 
                             "DivTaxo_Shann", "DivPhylo_Shann", "DivFctnal_Shann"),
                 variable.name = "explanatory",
                 value.name = "explanatory_value") %>% 
  melt(id.vars = c("plot", "Sylviculture", 
                             "explanatory", "explanatory_value"),
                 variable.name = "response",
                 value.name = "response_value") %>% 
 ggplot() +
 aes(x = explanatory_value, 
     y = response_value,
     fill = Sylviculture, col = Sylviculture) +
  geom_point(size = 2.1) +
  labs(x = "Explanatory value", y = "Response value") +
 theme_minimal()+
  geom_smooth(method = "lm") +
  ggpubr::stat_cor()+
  facet_grid(response ~ explanatory, scales = "free")

```
There are interactions between Sylviculture and Cedrela/soils/fire
And for between the other covariants ?

```{r Regressions with fire & Cedrela categorized}

#Split Cedrela and fire variables in classes :
covariantsdata_classes <- covariantsdata_plus1 %>% 
  mutate(Cedrela_classes = gtools::quantcut(prop_ced, 3)) %>% 
  mutate(Fire_classes = gtools::quantcut(prop_fire, 3)) 

write.csv2(covariantsdata_classes, "C:/Users/Utilisateur/Desktop/Stage M2 Côte d'Ivoire/covariantsdata_classes.csv")

covariantsdata_classesLongformat <- covariantsdata_classes %>% 
  select(plot, Sylviculture, Cedrela_classes, Fire_classes) %>% 
  left_join((covariantsdataLongformat))

# Under Cedrela proportion :
covariantsdata_classesLongformat %>%
  dcast(plot + Sylviculture + Cedrela_classes + Fire_classes ~ variable) %>% 
  melt(id.vars = c("plot", "Sylviculture", "Cedrela_classes", "Fire_classes", 
                             "DivTaxo_Shann", "DivPhylo_Shann", "DivFctnal_Shann"),
                 variable.name = "explanatory",
                 value.name = "explanatory_value") %>% 
  melt(id.vars = c("plot", "Sylviculture", "Cedrela_classes", "Fire_classes",
                             "explanatory", "explanatory_value"),
                 variable.name = "response",
                 value.name = "response_value") %>% 
 ggplot() +
 aes(x = explanatory_value, 
     y = response_value, col = Cedrela_classes) +
  geom_point(size = 2.1) +
  labs(x = "Explanatory value", y = "Response value") +
 theme_minimal()+
  geom_smooth(method = "lm") +
  ggpubr::stat_cor()+
  facet_grid(response ~ explanatory, scales = "free")

# Under Fire proportion :
covariantsdata_classesLongformat %>%
  dcast(plot + Sylviculture + Cedrela_classes + Fire_classes ~ variable) %>% 
  melt(id.vars = c("plot", "Sylviculture", "Cedrela_classes", "Fire_classes", 
                             "DivTaxo_Shann", "DivPhylo_Shann", "DivFctnal_Shann"),
                 variable.name = "explanatory",
                 value.name = "explanatory_value") %>% 
  melt(id.vars = c("plot", "Sylviculture", "Cedrela_classes", "Fire_classes",
                             "explanatory", "explanatory_value"),
                 variable.name = "response",
                 value.name = "response_value") %>% 
 ggplot() +
 aes(x = explanatory_value, 
     y = response_value, col = Fire_classes) +
  geom_point(size = 2.1) +
  labs(x = "Explanatory value", y = "Response value") +
 theme_minimal()+
  geom_smooth(method = "lm") +
  ggpubr::stat_cor()+
  facet_grid(response ~ explanatory, scales = "free")
 
```
There are interactions between Cedrela and soils/fire/sylviculture

* There is a positiv effect of fire & sylviculture on diversity when Cedrela proportion is low.
* Strong Cedrela proportion cancels fire  & sylviculture effects on diversity. 
* Cedrela intermediary-class allows higher diversity values.

same response of taxo & phylo