---
title: "Calculation of functional diversity indices"
author: "Vincyane Badouard"
date: "09/07/2020"
output: html_document
---


```{r Packages}

library(tidyverse) 
library(mice)
library(FD)
library(phylobase)
library(entropart)

```
https://daijiang.name/en/2014/05/11/functional-diversity-in-r/

0) Complete missing values with *MICE*
1) Take only traits
2) Compute species mean of each trait
3) Standardize all the traits (*scale()*) because their units are different
4) Obtain a dissimylarity matrix *for quanti & quali variables* : Gower method (*gowdis()*) 
5) Create a dendrogram (*hclust()*)
  - choose the agglomeration method to compute it (UPGMA) (ou tt faire et calculer leur Norme2 (pas corrélation cophénétique)
6) Test if it's utrametric & binary or force to be
7) Create a list of species composition table of all the plots (with only the species present in the traits base or not) 
8) Create a subset of this tree for each plot according to its species composition
9)Compute functional diversity indices


```{r Remove NA : No mesured species + MICE}

summary(Traitsbase)
names(Traitsbase)

Traitsbase <- Ourbase_MarcoRoots %>% 
filter(!(OldVernacular == "Ouossoupalie a fleurs mauves" & TreeID == "C")) %>% 
  select(-'RootAverageLenght', -'Filled', -'Operator')

methods(mice)
# methods used by this package are:
# PMM (Predictive Mean Matching)  – For numeric variables
# logreg(Logistic Regression) – For Binary Variables( with 2 levels)
# polyreg(Bayesian polytomous regression) – For Factor Variables (unordered levels >= 2)
# Proportional odds model (ordered, >= 2 levels)

# 2 variables with NA's :
# - Deciduousness -> 'polyreg' method
# - SPAD -> 'PMM' method



MICEinf <- mice(Traitsbase, maxit=100) #to use 100 iterations for each imputed dataset
MICEinf$imp$SPAD
MICEinf$imp$Deciduousness
Traitsbase_completed <- complete(MICEinf,5) #put these infered values in our dataset. We take the fifth estimation.

summary(Traitsbase_completed)

```


```{r Only species average traits}
names(Traitsbase)

Traitsbase_means <- Traitsbase_completed %>%
  select(genus,species,PetioleL,LeafThickness,LDMC,SLA,MeanDBH,meanWD,BarkTh,RDMC,RSL,RSA,Dispers,Deciduousness,SPAD) %>%
  unite(genus, species, col = "ScientificName", sep = " ", remove = T) %>%
  group_by(ScientificName) %>%
  mutate_if(is.numeric, mean) %>% 
  ungroup() %>% 
  distinct()
#13 traits, 48 species
```

```{r Obtain a dissimylarity matrix}

#Standardize numeric traits
Traitsbase_stand <- Traitsbase_means %>% 
  mutate_if(is.numeric,
            funs(as.vector(scale(.)))) %>% 
  column_to_rownames(., var = "ScientificName") #put species name in the column "rownames"

DistMatrix <- gowdis(Traitsbase_stand)
#Gower for quanti & quali variables

```


```{r Create a dendrogram (utrametric)}


FctDendrogram <-hclust(DistMatrix, method = "average")
# UPGMA : arithmetical average
plot(FctDendrogram, main="Functional dendrogram")
plot(FctDendrogram, hang=-1, main="Lien moyen") #hang : ?

require(clue)

UltrametricMatrix <- cophenetic(FctDendrogram)

#Norme 2
plot(DistMatrix,UltrametricMatrix,xlim=c(0,max(DistMatrix,UltrametricMatrix)), ylim=c(0,max(DistMatrix,UltrametricMatrix)),main="U-lien moyen vs D" )
abline(a=0, b=1, col = 3)
nc=round(cl_dissimilarity(DistMatrix, UltrametricMatrix, method = "spectral"),2)
text(0.43, 0.1, "norme 2 =")
text(0.5, 0.1, nc)

class(FctDendrogram) #it is an object of "hclust" class
FctDendrogram <- as.phylo(FctDendrogram)#transform in "phylo class" for some following functions
is.ultrametric(FctDendrogram) #It's an ultrametric tree
is.binary.tree(FctDendrogram) #It's a binary tree
plot(FctDendrogram, main="Functional utrametric tree")

FctDendrogram_phylo4 <- as(FctDendrogram, "phylo4")#transform in "phylo4 class" for "subset" functions

save(FctDendrogram, file = "FctDendrogram.Rdata")
# load("FctDendrogram.Rdata")
```

```{r a tree for each plot}
# list with a table for each plot : 
datafctnaltree <- Data_Tene %>% #species composition table for each plot
  select(genus, species, family, plot) %>% 
  distinct() %>% 
  filter(!is.na(genus)) %>% #remove non-identified species
  filter(!(genus == "Cedrela")) %>% #Remove Cedrela of the response variable because it will be a covariant
  unite(genus, species, col = "ScientificName", sep = " ", remove = F) %>% 
  unite(genus, species, col = "Scientific_Name", sep = " ", remove = F)

#create a list with a table for each plot :
datafctnaltreelist <- split(datafctnaltree, datafctnaltree$plot)

datafctnaltreelist <- lapply(datafctnaltreelist, function(element) column_to_rownames(element, var = "Scientific_Name"))  #put species name in the column "rownames" for 'subset' function


#a dendrogram for each plot :
fctnalTreeperplot <- lapply(datafctnaltreelist, function(element) subset(FctDendrogram_phylo4, tips.include = element$ScientificName))
save(fctnalTreeperplot, file = "fctnalTreeperplot.Rdata")
# load("fctnalTreeperplot.Rdata")


```

```{r Functional diversity indices computation}

# 'bcPhyloDiversity' function necessites a 'NAMED VECTOR' as species abundances numeric vector (non specified in R help's function) :

# Species abundances numeric NAMED vector
Namedvector_fctn <- lapply(abondplotslist,
                      function(element){
                         Namedvector_fctn <- element$N   # operation1
                         names(Namedvector_fctn) <- element$ScientificName # operation2
                         return(Namedvector_fctn) # préciser ce que la fonction doit renvoyer
})

str(Namedvector_fctn)

fctnalTreeperplot <- lapply(fctnalTreeperplot, function(element) as(element, "phylo"))#transform in "phylo class" to compute diversity indices

DivFctnal_Shann <- mapply(function(X,Y) {
  bcPhyloDiversity(Ns = X,
                   Tree = Y,
                   q = 1, Normalize = T,
                   Correction = "Best", CheckArguments = T)
  }, X=Namedvector_fctn, Y=fctnalTreeperplot, SIMPLIFY = F) #SIMPLIFY = F : laisse la structure initiale de la liste (par plot dans ce cas)

DivFctnal_Shann_values <- lapply(DivFctnal_Shann, getElement, "Total")#extraction de la valeur (Total) de l'indice grace à la fonction "getElement"(='$')

```

```{r Put values in a table}
Diversitytable_fctnalshan <- DivFctnal_Shann_values %>% 
  as_tibble(.) %>% 
  t() %>% #transpose romws & columns
  as.tibble(.) %>%
  rownames_to_column(var= "plot") %>% 
  rename(DivFctnal_Shann=V1)
```
